Can a High Color Contrast Touch Interface Increase User Reaction Time when Using a Smart Phone Web Based Application.

### 2015-11-17 Paper.tex
Updates to final paper

\section{Introduction}
The main goal with this paper is to evaluate if color contrast has a significant impact on usability in web based smart-phone applications. Consumer touch screen devices such as smart-phones has rapidly increased in amount and availability recent years. The touch screen technology have made great advances ~\cite{jennings2013touch} and it is more frequently used as a way of receiving user input. When this technology moves to a broader audience, higher demands on usability needs to be set~\cite{gong2004guidelines}. To further explore how usability can be improved using touch screen technology we are in this paper investigating if reaction time (referred to as RT\footnote{Reaction Time: the time it takes for a user to react on instructions} in this paper) and user input errors (referred to as IE\footnote{Input Errors: the number of user input errors when interacting with a system} in this paper) can reach better results by increasing color contrast within the interface. Earlier work done in this field are exploring how we perceive different color contrast and how it can affect our reading performance~\cite{wu2003improving}. It has also been proved that Chromaticity, Contrast, and Cone Opponency in color space can affect RTs~\cite{mckeefry2003simple}. Many best practices for Mobile development also shows that "requirements for sufficient color contrast"~\cite{marcus2013design} must be meet to best exaggerate the content. But still the question remains, if a high color contrast interface can increase the the RT and decrease user IE.

High RTs and low user IE are especially desirable when designing user interfaces for situations with high demands on quick user input and low error tolerance such as emergency situations. The results of this paper can be used by any designer or developer.

\section{Method}
In order to be able to test if color contrast has an substantial impact on the RT and user IE of a touch user interface we have designed a simple web application for an Android smart-phone. The application exists in two versions, one with low color contrast (referred to as LC in this paper) and one with high color contrast (referred to as HC in this paper) as can be seen in Fig~\ref{fig:application}.

@@

We designed the application by following design guidelines given here~\cite{hoober2011designing}~\cite{johnson2013designing}~\cite{gong2004guidelines}~\cite{norman2013design}.

\subsection{Designing the A/B test}
We conducted an A/B test to measure the differences in RT and IE between the two versions. An A/B test is a commonly used method by developers and designers to test differences in performance between applications when the differentiated factors are known~\cite{johnson2013designing}. As in this case our known factor is the contrast difference.

The application consists of 4 buttons\footnote{On screen button}, each button representing a function. To fraction each button from the others we put a unique shape in each button. The shapes we used are a, square, circle, triangle and a rhomb~\ref{fig:application}.

@@

\subsection{Evaluation of the A/B test}
Our test data received from the tests consisted of UNIX timestamps\footnote{\url{http://www.unixtimestamp.com/}} in milliseconds. Each time the user interacted with the application it registered which button that was pressed and saved it together with a UNIX timestamp. Equally, every time the instruction program showed a new symbol it logged that together with a UNIX timestamp. As a result we received two data sets of symbols with corresponding timestamps for each test person, resulting in a total of approximately 400 records. Although the data were hard to interpret due to its compact look, a python script were written to clean it up. We calculated each difference (RT) in time between when the instruction were given and when user interaction registered. Out of this we could calculate the mean RT for both groups LC and HC. The calculation of RT did not take in consideration if user input were a no match. We still considered it as an reaction.

>>>>

\section{Introduction}
The main goal with this paper is to evaluate if color contrast has a significant impact on usability in web based smart-phone applications. Consumer touch screen devices such as smart-phones have rapidly increased in amount and availability recent years. The touch screen technology have made great advances~\cite{jennings2013touch} and it is frequently used as a way of receiving user input. When this technology moves to a broader audience, higher demands on usability needs to be set~\cite{gong2004guidelines}. To further explore how usability can be improved using touch screen technology we have in this paper investigated if reaction time (the time it takes for a user to react on instructions, further referred to as RT in this paper) and user input errors (the number of user input errors when interacting with a system, further referred to as UIE in this paper) can reach better results by increasing color contrast within the interface. Better results are assumed as faster RTs and less number of UIEs.

Earlier studies completed in this field has explored how we perceive different color contrast and how it can affect our reading performance~\cite{wu2003improving}. It has also been proved that chromaticity, contrast, and cone opponency in color space can affect RTs~\cite{mckeefry2003simple}. Many best practices for Mobile development have also shown that "requirements for sufficient color contrast"~\cite{marcus2013design} must be meet to best exaggerate the content. But still the question remains, if a high color contrast interface can increase the the RT and decrease user IE.

High RTs and low user IE are especially desirable when designing user interfaces for situations with high demands on quick user input and low error tolerance such as in emergency situations. The results of this paper can be used by any designer or developer.

\section{Method}
In order to be able to test if color contrast has an substantial impact on the RT and user IE of a touch user interface we have designed a simple web application for an Android smart-phone. The application exists in two versions, one with low color contrast (referred to as LCC in this paper) and one with high color contrast (referred to as HCC in this paper) as can be seen in Fig~\ref{fig:application}.

@@

We designed the application by following design guidelines given here~\cite{hoober2011designing}~\cite{johnson2013designing}~\cite{gong2004guidelines}~\cite{norman2013design}. These guidelines can be seen as a set of rules to follow when designing any kind of user interface.

\subsection{Designing the A/B test}
We conducted an A/B test to measure the differences in RT and UIE between the two versions. An A/B test is a commonly used method by developers and designers to test differences in performance between applications when the differentiated factors are known~\cite{johnson2013designing}. Where in this case our known factor is the contrast difference.

The application consists of 4 buttons\footnote{On screen button}, each button representing a function. To differentiate each button from the others we put a unique shape in each button. The shapes we used where a, square, circle, triangle and a rhomb (Fig~\ref{fig:application}).

@@

\subsection{Evaluation of the A/B test}
Our test data received from the tests consisted of UNIX timestamps\footnote{\url{http://www.unixtimestamp.com/}} in milliseconds. Each time the user interacted with the application it registered which button that was pressed and saved it together with a UNIX timestamp. Equally, every time the instruction program showed a new symbol it logged that together with a UNIX timestamp. As a result we received two data sets of symbols with corresponding timestamps for each test person, resulting in a total of approximately 400 records. Although the data were hard to interpret due to its compact look, a python script were written to clean it up. We calculated each difference (RT) in time between when the instruction were given and when user interaction registered. Out of this we could calculate the mean RT for both groups LCC and HCC. The calculation of RT did not take in consideration if user input were a no match. We still considered it as an reaction.
